---
title: "Week 2"
author: "Me"
date: "12/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#MySql

```{r}
install.packages('RMySQL')
library(RMySQL)
```

connect database
```{r}
ucsc_db = dbConnect(MySQL(), user = 'genome', host = 'genome-mysql.cse.ucsc.edu')
result = dbGetQuery(ucsc_db, 'show databases;')
```

connecting to hg19
```{r}
hg19 = dbConnect(MySQL(), user = 'genome', db = 'hg19', host = 'genome-mysql.cse.ucsc.edu')
all_tables = dbListTables(hg19)
length(all_tables)
head(all_tables)
```

Get dimensions of a specific table
```{r}
dbListFields(hg19, 'affyU133Plus2')
```

```{r}
dbGetQuery(hg19, 'SELECT COUNT(*) FROM affyU133Plus2')
```

Read data from table
```{r}
affy_data = dbReadTable(hg19, 'affyU133Plus2')
```

Select specific subset

```{r}
query = dbSendQuery(hg19, 'SELECT * FROM affyU133Plus2
                              WHERE misMatches BETWEEN 1 AND 3')

affy_mis = fetch(query)
quantile(affy_mis$misMatches)
```

```{r}
affy_mis_small = fetch(query, n = 10)
dbClearResult(query) #clear query from remote server
affy_mis_small
```

Close connection
```{r}
dbDisconnect(hg19)
dbDisconnect(ucsc_db)
```

# HDF5

For large datasets
Hierarchical data format

Groups containing zero or more datasets and metadata
- have a group header with group name and list of attributes
- have a group symbol table with list of objects in a group.

Datasets multidimensional array of data elements with metadata.
- have a header with name, datatype, datastorage, storage layout

```{r}
install.packages("BiocManager")
library(BiocManager)
BiocManager::install("rhdf5")
library(rhdf5)
```
```{r}
created = h5createFile('example.h5')
```
Create groups
```{r}
created = h5createGroup('example.h5', 'foo')
created = h5createGroup('example.h5', 'baa')
created = h5createGroup('example.h5', 'foo/foobaa') #subgroup

#list files
h5ls('example.h5')
```

Write to groups
```{r}
A = matrix(1:10, nr = 5, nc = 2)
h5write(A, 'example.h5', 'foo/A')

B = array(seq(.1, 2, by = .1), dim = c(5,2,2))
attr(B, 'scale') = 'liter' #add attribute

h5write(B, 'example.h5', 'foo/foobaa/B')
h5ls('example.h5')
```

read data
```{r}
h5read('example.h5', 'foo/A')
```

Writing and reading chunks
```{r}
h5write(c(12,13,14), 'example.h5','foo/A', index = list(1:3,1))
h5read('example.h5','foo/A')
```

# Reading data from the web

```{r}
con = url('https://scholar.google.com/citations?user=HI-I6C0AAAAJ&hl=en')
html_code = readLines(con)
close(con)
html_code
```

Parsing with XML
```{r}
library(XML)
url = 'http://scholar.google.com/citations?user=HI-I6C0AAAAJ&hl=en'
html = htmlTreeParse(url, useInternal = T)
```

Use get from httr
```{r}
library(httr)
html = GET(url) #
content = content(html, as = 'text') #extract content from html page
parsed_html = htmlParse(content, asText = T) #to parse out text to look like html
xpathSApply(parsed_html, '//title', xmlValue)
```

Accessing websites with password
```{r}
pg1 = GET('http://httpbin.org/basic-auth/user/passwd')
```

```{r}
pg2 = GET('http://httpbin.org/basic-auth/user/passwd', 
          authenticate('user','passwd'))
pg2
```

Using handles
```{r}
google = handle('http://google.com')
pg1 = GET(handle = google, path = '/')
pg2 = GET(handle = google, path = 'search')
```


##Reading from APIs

```{r}
my_app = oauth_app('twitter', key = 'consumerkeyhere', secret = 'consumersecrethere')
sig = sign_oauth1.0(my_app. token = 'token here', token_secret = 'token secret here')
home_tl = GET('https://api.twitter.com/1.1/statuses/home_timeline.json', sig) #will return json data
```


```{r}
json1 = content(home_tl)
json2 = jsonlite::fromJSON(tojson(json1)) #json structure - structured object back to json - fromJson to create dataframe
json[1,1:4]
```


httr allows GET, POST, PUT, DELETE if you are authorized
Most modern APIs use oauth
httr works well with facebook, google, twitter, github, etc.

## Readign from other sources

Read images
Read from other database
Read GIS data
Reading Music data


#Quiz
```{r}
library(httr)
library(jsonlite)

# 1. Find OAuth settings for github:
#    http://developer.github.com/v3/oauth/
oauth_endpoints("github")

# 2. To make your own application, register at
#    https://github.com/settings/developers. Use any URL for the homepage URL
#    (http://github.com is fine) and  http://localhost:1410 as the callback url
#
#    Replace your key and secret below.
myapp <- oauth_app("Coursera_Getting_Cleaning_Data",
  key = "key",
  secret = "secret"
)

# 3. Get OAuth credentials
github_token <- oauth2.0_token(oauth_endpoints("github"), myapp)

# 4. Use API
gtoken <- config(token = github_token)
req <- GET("https://api.github.com/users/jtleek/repos", gtoken)
stop_for_status(req)
content(req)

# OR:
req <- with_config(gtoken, GET("https://api.github.com/users/jtleek/repos"))
stop_for_status(req)
content(req)

```

```{r}

json = content(req)
data = fromJSON(toJSON(json))

library(tidyverse)
data %>%
      filter(str_detect(full_name,'datasharing')) %>%
      select(created_at) %>%
      unlist(.$created_at)
```

```{r}
url = 'http://biostat.jhsph.edu/~jleek/contact.html'
conn = url(url)
html_lines = readLines(conn)
close(conn)

for (i in c(10, 20, 30, 100)){
      print(nchar(html_lines[i]))
}
```

```{r}
url = 'https://d396qusza40orc.cloudfront.net/getdata%2Fwksst8110.for'
lines = readLines(url, n = 10)

#got this from stackoverflow
data = read.fwf(file = url,  skip= 4,
  widths=c(12, 7, 4, 9, 4, 9, 4, 9, 4)) 

#sum of 4th column
data %>%
      select(V4) %>%
      summarise(sum = sum(V4))
```


