---
title: "Week 4"
author: "Me"
date: "12/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Editing text variables

tolower() - converts to lower
toupper()
strsplit(x, 'character') 
sub('_', '', names(reviews),)
gsub - replace mutiple instances
grep() - finds index that has search string. Value = T means returns the value where it appears
grepl() - returns a vector of true when string appears
stringr fucntions
 - nchar()  - number of chars
 - substr() - take out part of strings
 - paste() - Paste strings
 - paste0() - no space
 - str_trim() - removes whitespace
 
 
 #Regular expressions

- combination of literals and metacharacters. 
- Literals - match exactly
- use meta characters to specify more general search terms
- metacharacters
      - ^I - starts with I
      - morning$ - ends with morning
      - [list] [of] [characters]. [Bb][Uu][Ss][Hh]
      - ^[Ii] am - match lower of upper I followed by am.
      - range of characters ^[0-9][a-zA-z] all letters.
      - [^?.]$ not ends with ? or . '^' here is NOT
      - . is any character
      - | combine flood|fire - either flood or fire. Can combine any characters
      - ^[Gg]ood|[Bb]ad
      - ^([Gg]ood|[Bb]ad) - constrains alternatives
      - [Gg]eorge ([Ww]\.)? [Bb]ush - ? is optional
      - * repeat any number of times. 
      - + means at least one.
      - [0-9]+ (.*) [0-9]+ - at least 1 num followed by any char followed by at least 1 num.
      - {} minimum and max number of expressions
      - [Bb]ush ( +[^ ]+ +){1,5} debate. Bush + (one space + not space + one space ){at least 5x} debate
      - m,n = at least m but not more than n
      - m means exactly m matches
      - m at least m matches
      - () use to remember text matched. 
      - +([a-zA-z]+) +\1 + = space word space same word
      - * is greedy so it matches longest possible string
      - ^s(.*)s - starts with s and ends with s
      - ^s(.*?)s$ - less greedy           


#working with dates
```{r}
d1 = date()
class(d1)
```
```{r}
d2 = Sys.Date()
class(d2)
```

Formatting dates
```{r}
format(d2, '%A %b %d')
```
```{r}
x = c('1jan1960', '2jan1960', '31mar1960', '30jul1960')
z = as.Date(x, '%d%b%Y')
z[1] - z[2]
as.numeric(z[1] - z[2])
```
```{r}
weekdays(d2)
months(d2)
julian(d2)
```

Lubridate
```{r}
library(lubridate)
ymd('20140108')
mdy('08042013')
ymd_hms('2011-08-03 10:15:03')
```

#QUiz
```{r}
library(tidyverse)
```

```{r}
data = read_csv('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv')
sapply(names(data), strsplit, "wgtp")[123]
```

```{r}
data = read_csv('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv', skip = 4) %>%
   select(c(1,2,4,5)) %>%
   rename(country_code = X1, rank = X2, country = X4, gdp = X5)

#wrong answer
data %>%
   mutate(gdp = parse_number(gdp)) %>%
   summarise(mean(gdp, na.rm = T))

#must filter with all cases
data %>%
   filter(complete.cases(.)) %>%
   mutate(gdp = as.numeric(str_replace_all(gdp, ',',''))) %>%
   summarise(mean(gdp))
   
```

Get all united
```{r}
data %>%
   filter(str_detect(country, '^United'))
```

```{r}
gdp = data
educational = read_csv('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv') %>%
   janitor::clean_names()

merged = merge(gdp, educational, by = 'country_code')
merged %>%
   filter(str_detect(special_notes,'(Fiscal year)(.*)(June)'))
```

```{r}
library(quantmod)
amzn = getSymbols("AMZN",auto.assign=FALSE)
sampleTimes = index(amzn)
st = as_tibble(sampleTimes)
st %>%
   mutate(date = ymd(value)) %>%
   filter(year(date) == 2012) %>%
   mutate(day = wday(date, label = T)) %>%
   filter(day == 'Mon')
```















