---
title: "Week 3"
author: "Me"
date: "12/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Subsetting and sorting
```{r}
set.seed(13435)
X = data.frame('var1' = sample(1:5), 'var2'= sample(6:10), var3 = sample(11:15))
X[,1]
```
```{r}
X[X$var1<3, ]
```

Which does not return NA
```{r}
X[which(X$var2 > 8),]
```
Sorting
```{r}
sort(X$var1)
sort(X$var1, decreasing = T)
```

Order
```{r}
X[order(X$var1),]
X[order(X$var1, X$var3),]
```

## Summarizing data


Functions
head
tail
summary
str
quantile(na.rm = T, probs = c(.5, .75, .9))
table(x, useNA = 'ifany')
sum(is.na())
any(is.na())
all()
colSums()
all(colsums(is.na(data)) = 0) #TRUe if no missing
table(var %In% c('A','B'))

Crosstabs
```{r}
data("UCBAdmissions")
DF = as.data.frame(UCBAdmissions)
xt = xtabs(Freq ~ Gender + Admit, data = DF)
xt
```

Flat table
```{r}
data("warpbreaks")
warpbreaks$replicate = rep(1:9, len = 54)
xt = xtabs(breaks ~., data = warpbreaks) #value is breaks
ftable(xt)
```

#Creating new variables

Common variables to create
- missingness indicator
- cutting up variables
- transformations

Sequences
```{r}
seq(1,10, by = 2)
```
Creating variable using ifelse
```{r}
ifelse(condition, TRUE, FALSE)
```

Creating categorical variables using cut
```{r}
cut(variable, breaks = value)
```

Easier cutting using cut2 on Hmisc library

Creating factor variables using factor()

Levels of factor variables
```{r}
yesno = sample(c('yes','no'), size = 10, replace = T)
yesnofac = factor(yesno, levels = c('yes','no'))
relevel(yesnofac, ref = 'yes')
```
Common transforms
abs(x)
sqrt(x)
ceiling(x) 
floor(x)
round(x, digits = n)
signif(x, digits = n)
cos(x), sin(x) etc
log(x)
log2(x), log10(x)
exp(x)


#Quiz

```{r}
library(tidyverse)
```

```{r}
data = read_csv('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv')
head(data)
agricultureLogical = data$ACR == 3 & data$AGS == 6
which(agricultureLogical)
```

```{r}
library(jpeg)
data = readJPEG('getdata_jeff.jpg', native = T)
quantile(data, probs = c(.3,.8))
```

```{r}
gdp = read_csv('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv', skip = 3) %>%
      select(c(1,2,4,5)) %>%
      janitor::clean_names() %>%
      rename(country_code = x1)

educational = read_csv('https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv') %>%
      janitor::clean_names()


merge(gdp, educational, by = c('country_code' = 'country_code')) %>%
      filter(!is.na(ranking)) %>%
      mutate(ranking = as.numeric(ranking)) %>%
      arrange(desc(ranking)) %>%
      slice(n = 13)
```

```{r}
merge(gdp, educational, by = c('country_code' = 'country_code')) %>%
      filter(!is.na(ranking)) %>%
      filter(income_group =='High income: nonOECD' | income_group == 'High income: OECD') %>%
      mutate(ranking = as.numeric(ranking)) %>%
      group_by(income_group) %>%
      summarise(ranking = mean(ranking))
      
```
```{r}
merged = merge(gdp, educational, by = c('country_code' = 'country_code')) %>%
      filter(!is.na(ranking)) %>%
      mutate(ranking = as.numeric(ranking))

quantile_ranking = quantile(merged$ranking, probs = seq(0,1, .2), na.rm = T)

merged %>%
      mutate(quantile_ranking = cut(.$ranking, breaks = quantile_ranking)) %>%
      filter(income_group=='Lower middle income') %>%
      group_by(quantile_ranking) %>%
      count() %>%
      arrange(quantile_ranking)
```




















